Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-21 article.tex
80 char break
wording + comment

\item \textbf{GENIE3} \citep{huynhthu2010inferring} translates the inference
of the network of $p$ neurons into $p$ supervised learning tasks. For each
task, the fluorescence of a neuron is predicted from the fluorescence of all
other neurons using a tree based ensemble such as random forest
\citep{breiman2001random}. The tree ensemble associates a variable importance
score \citep{louppe2013understanding} to each neuron used in the predictions,
which is used as a degree of associate. The following parameters are used
$10000$ trees, $\log_2{p}$ variables are randomly selected to split tree nodes
and the maximal tree depth is constrained to 1, 2 or 3. \end{itemize}

@@

Figure \ref{fig:curves} shows the ROC curve  and precision-recall curve
for correlation, GTE and partial correlation on the \textit{normal-2}
dataset. The partial correlation outperforms other state-of-the-art methods
by at least factor of 2 in AUPRC reaching $0.347$ and an increased of $0.05$ in
AUROC. Two reasons might explain such differences: (i) among
those methods only the partial correlation makes the distinction between
direct and indirect connections, (ii) the simulated network from the
\citep{stetter2012model} generator or neural networks in general could be
reduced to a linear model after pre-processing. Those results is obviously
generalized over the other datasets see Appendix \textcolor{red}{REF}.

@@

As observed on Figure \ref{fig:curves}, the GENIE3 performance
increases with deeper trees while being still computationally feasible. The
reason is twofold. First, the deeper the tree, the larger is the conditioning.
\textcolor{red}{AJ: conditioning?}
As it was observed in the comparison between correlation and partial
correlation, it screens indirect links in the inferred network.
\textcolor{red}{AJ: I don't understand the link with GENIE3}
Second, the
number of small trees need to be much higher than deep trees in order to
observe - and evaluate - all pairs.  However, as suggested by
\cite{louppe2014understanding}, limiting the maximal depth of trees may help
to improve the variable importance computing, i.e.
inferring the degree of association. \textcolor{red}{AJ: there is no explanation
of why it would be better apart for computational reason.} Therefore a trade-off
 of the maximal depth may give the best of both worlds. \textcolor{red}{AJ:
 Which tradeoff? Which worlds?}

% Opt: obviously when parameters are tuned/optimized, it's slightly better -->
% see appendix for a full description of the method

% \begin{table}
% \small
% \centering
% \begin{tabular}{@{}l *{8}{c}@{}}
%   & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
% \cmidrule(r){2-5}\cmidrule(r){6-9}
% Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
% \hline \hline
% % Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
% GENIE3 (depth = 1)     &  & & & & & & & \\
% GENIE3 (depth = 2)     &  & & & & & & & \\
% GENIE3 (depth = 3)     &  & & & & & & & \\
% GTE                    &  & & & & & & & \\
% % Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
% \end{tabular}
% \caption{TODO (network are undirected)}
% \label{tab:tab2}
% \end{table}

@@

If the inferred network corresponds to the undirected network, it would
already yield high AUROC score for instance $\approx 0.996$ on
\textit{normal-1}. Infering edge directivity is likely to be counter-
productive or yield small gain. By contrast, the same experiment with AUPRC
would yield to a score of $\approx 0.789$ on \textit{normal-1}. Thus inferring
correctly directivity would be more rewarding in AUPRC term.

@@

The area under the ROC curve is not consistent metric
with respect to the problem statement. A clear hierarchy is also difficult to
derive based on ROC curve of Figure \ref{fig:roc_curve}. We advice to avoid the AUROC
for the assessment of network inference method. An alternative metric would
be the area under the precision-recall curve (AUPRC). Those
conclusions are in line with \cite{schrynemackers2013protocols}.

>>>>

\item \textbf{GENIE3} \citep{huynhthu2010inferring} translates the inference
of the network of $p$ neurons into $p$ supervised learning tasks. For each
task, the fluorescence of a neuron is predicted from the fluorescence of all
other neurons using a tree based ensemble such as random forest
\citep{breiman2001random}. The tree ensemble associates a variable importance
score \citep{louppe2013understanding} to each neuron used in the predictions,
which is used as a degree of associate. The following parameters are used
$10000$ trees, $\log_2{p}$ variables are randomly selected to split tree nodes
and the maximal tree depth is constrained to 1, 2 or 3. \textcolor{red}{AJ Which
depth is represented on the graph?}
\end{itemize}

@@

Figure \ref{fig:curves} shows the ROC curve  and precision-recall curve
for correlation, GTE and partial correlation on the \textit{normal-2}
dataset. The partial correlation outperforms other state-of-the-art methods
by at least factor of 2 in AUPRC reaching $0.347$ and an increased of $0.03$ in
AUROC. Averaging Partial correlation matrix over a wide of parameters
further improves AUPRC to $0.389$ and improve AUROC by $0.009$.
Two reasons might explain such differences: (i) only the partial
correlation-based methods and GENI3 at lesser extent distinguish
direct links from indirect one, (ii) simulated networks from the
\citep{stetter2012model} generator or neural networks in general could be
reduced to a linear model after pre-processing. Those results
generalizes over the other datasets see Appendix \textcolor{red}{REF}.

@@

\textcolor{red}{Only one GENI3, thus it can't be observed.}
As observed on Figure \ref{fig:curves}, the GENIE3 performance increases with
deeper trees while being still computationally feasible. The reason is twofold.
First, the deeper the tree, the larger is the conditioning. \textcolor{red}{AJ:
conditioning?} As it was observed in the comparison between correlation and
partial correlation, it screens indirect links in the inferred network.
\textcolor{red}{AJ: I don't understand the link with GENIE3} Second, the number
of small trees need to be much higher than deep trees in order to observe - and
evaluate - all pairs.  However, as suggested by \cite{louppe2014understanding},
limiting the maximal depth of trees may help to improve the variable importance
computing, i.e. inferring the degree of association. \textcolor{red}{AJ: there
is no explanation of why it would be better apart for computational reason.}
Therefore a trade-off  of the maximal depth may give the best of both worlds.
\textcolor{red}{AJ:  Which tradeoff? Which worlds?}

@@

If the inferred network corresponds to the undirected network, it would
already yield high AUROC score for instance $\approx 0.996$ on
\textit{normal-1}. Infering edge directivity is likely to be counter-
productive or yield small gain. By contrast, the same experiment with AUPRC
would yield a score of $\approx 0.789$ on \textit{normal-1}. Thus inferring
correctly directivity would be more rewarding in AUPRC term.

@@

The area under the ROC curve is not a consistent metric with respect to the
problem statement. A clear hierarchy is also difficult to derive from ROC
curves as shown on Figure \ref{fig:roc_curve}. We advice to avoid the AUROC for
the assessment of network inference method. An alternative would be to assess
network inference using the area under the precision-recall curve (AUPRC).
Those conclusions are in line with \cite{schrynemackers2013protocols}.
